Recommendation Generation for Performance Improvement by using Cross-Organizational Process Mining

### 2015-09-08 5_results_discussions/results_discussions.tex
no message

Approach in this study is an aggregation of various methods and they are significantly different from each other in their mathematical background. Therefore, instead of a global evaluation metric for the complete methodology, each stage will be evaluated within its evaluation metrics. Since these stages are executed sequentially, it is important for each stage to perform well enough to yield a successful outcome. Evaluation metrics for each stage are as follows:

\begin{description}
	\item[Process Model Mining] Performance of process mining stage is measured by \textit{Fitness} and \textit{Appropriateness} which are defined in \cite{rozinat2008conformance}.
	\item[Performance Indicator Analysis] In replay phase, \textit{alignment costs} \cite{van2012replaying}  are compared with process model mining metrics. In clustering phase, within-SSE analysis is undertaken to decide on the number of clusters.
	\item[Mismatch Pattern Analysis] In this stage, number of mismatch patterns found are compared with the \textit{graph-edit similarity} \cite{dijkman2011similarity} of process models.
	\item[Recommendation Generation] In recommendation generation, different threshold values are tried to check how many mismatch patterns are generated for organizations and how they could be used for focused analysis.
\end{description}

Cross-organizational mining aims to find cross-correlation of workflows and activities in different organizations and this yields the necessity of organizations that do the same main activity with comparable process flows. Considering these characteristics, there are few dataset available in the literature that are well-structured, documented and valuable. In this thesis study, one synthetic and one real-life event log datasets are presented and used in the following sections to evaluate the performance of the proposed methodology.

\begin{description}
  \item[Loan Application Process \cite{loan-app-data}] This synthetically created dataset consists of event log variants of a simple loan application in a financial institute. This dataset includes artificial event logs of 4 variants where each variant includes different sets of approaches such as parallelism, choices and sequential tasks. These event logs are used to test different approaches of discovering a configurable process model from a collection of event logs \cite{buijs2014flexible}.
  \item[Environmental Permit Application Process \cite{coselog-data}] This dataset originates from the "Configurable Services for Local Governments (CoSeLoG)" project \cite{van2011business} which investigates the similarities and dissimilarities between several processes of different municipalities in Netherlands. Dataset contains records of receiving phase for the building permit application process in 5 municipalities, which are comparable since activity labels in the different event logs refer to the same activities performed in five municipalities. This dataset is also mentioned in the literature as \textit{"Processing applications for building and/or environmental permits (Wet Algemene Bepalingen omgevingsrecht (WABO) in Dutch)"}.
\end{description}

>>>>

Approach in this study is an aggregation of various methods and they are significantly different from each other in their mathematical background. Therefore, instead of a global evaluation metric for the complete methodology, each stage is evaluated within its evaluation metrics which are as follows:
\begin{description}
	\item[Process Model Mining] Performance of process mining stage is measured by \textit{fitness} and \textit{appropriateness} which are defined in \cite{rozinat2008conformance}.
	\item[Performance Indicator Analysis] In replay phase, \textit{alignment costs} \cite{van2012replaying} are compared with process model mining metrics. In clustering phase, \textit{within-SSE} analysis is undertaken to decide on the number of clusters.
	\item[Mismatch Pattern Analysis] In this stage, number of mismatch patterns found are compared with the \textit{graph-edit similarity} \cite{dijkman2011similarity} of process models.
	\item[Recommendation Generation] In recommendation generation, different threshold values are tried to check how many mismatch patterns are generated for organizations and how they could be used for focused analysis.
\end{description}

\input{5_results_discussions/loan-application-process/loan-application-process.tex}
\input{5_results_discussions/coselog-wabo/coselog-wabo.tex}

\subsection{Discussions}
\label{subsec:discussions}
When the evaluation of the stages for \textit{Loan Application Process} and \textit{Environmental Permit Application Process} datasets are gathered together, the following results can be expressed:

\begin{itemize}
	\item Process mining stage of the proposed methodology can mine the process models with high fitness appropriateness levels.
	\item For the successfully mined models with high fitness values, replay and performance indicator calculation stage works seamlessly as expected. With this step, average and standard deviation time between each activity can be measured for each organization. Number of these metrics are quadratic to the number of activities in each organization's process model and difficult to analyze with a cross comparison.
	\item Internal measure of clusters indicates that the organizations can be clustered according to their performance indicators which yields a collective approach of organizations for their subprocesses. In other words, organizations are divided into clusters which shows that they can be grouped based on how well they are executing.
	\item Mismatch analysis spots the differences between process models in coherence with structural similarity of them. This indicates that the idea of using mismatch patterns to reveal differences between process models is a feasible approach since its results are comparable to the similarity metrics of process models in the literature.
	\item Recommendation generation aims to gather all generated information in this study to help focusing on the potentially important mismatch patterns for performance improvement. When the number of mismatch patterns with and without performance clusterings are checked, it shows that in a small dataset performance clustering lists 3 times less number of differences in \textit{Loan Application Example} dataset. When it is impossible to locate mismatch patterns manually like in \textit{Environmental Permit Application Process}, performance clustering spots 100 times less number of differences. This difference helps user to focus on the differences with a potential performance improvement which is one of the aims in this thesis study.
	\item Although each step of methodology can be counted as successful based on their evaluation metrics, mismatch patterns recommended at the end of methodology can yield important observations as well as being irrelevant and infeasible. Since this decision is based on the business environment of organizations, evaluation of the quality of recommendations for business usefulness requires domain expertise. However, an example recommendation can be presented to provide an insight. In the analysis of \textit{Loan Application Process}, Variant \#3 performs worse 27 \%  on average time and 12 \% on standard deviation time between activities "Calculate Capacity" and "Accept". When the mismatch patterns for these performance indicators are checked the following ones can be mentioned:
		\begin{itemize}
		\item "Check Credit" is a \textit{Refined Activity} of with "Check System (50 \%)"; "Check Paper Archive (42 \%)"; "Send Credit Check Request (32 \%)"; "Process Credit Check Reply (31 \%)" where the corresponding similarity values provided in parentheses.
		\item "Calculate Capacity" is a \textit{Different Moments in Processes} which have different previous activities in clusters. 
		\end{itemize}
	When these example mismatch patterns are checked, removing "Check Credit" activity and putting other activities instead of it might be the cause of performance improvement. With the same approach, putting "Calculate Capacity" on different orders in processes can effect the average and variance of time between activities. These mismatch patterns are also visualized on process model of Variant \#3 and a variant from other cluster in Figure~\ref{fig:loan-recommendation-visualization}. In the process models, refined activities of "Check Credit" and different positions of "Calculate Capacity" are indicated. 
		\begin{figure}
			\centering
			\includegraphics[width=\textwidth]{5_results_discussions/loan-application-process/recommendation-visualization}
			\caption{Visualization of example recommendation for Loan Application Process dataset}
		  \label{fig:loan-recommendation-visualization}
		\end{figure}
\end{itemize} % end of discussions

### 2015-09-08 6_conclusion_future_work/conclusion_future_work.tex
no message

\section{Conclusion}

In this study, a new approach is proposed and tested for generating recommendations using cross-organizational process mining for process performance improvement. Cross-organizational process mining is applied with the idea of unsupervised learning where predictor variables related to performances of organizations are used in an environment where processes are executed on several organizations. Results show that it is possible to use cross-organizational process mining and mismatch patterns for performance improvement recommendations. Process mining is a large-spectrum field where different set of activities and approaches are gathered together to discover, monitor and improve processes. In this  study, a four-stage solution is presented and their performances are explained. In addition, proposed methodology is developed as extensible and configurable set of plugins in ProM framework \cite{verbeek2010prom} and published as open-source. This makes the methodology open to include new process mining methods, mismatch patterns and clustering approaches as well as testing with different datasets.

\section{Future Work}

For the approach proposed in this thesis study, the following issues can be listed as pointers to future work:
\begin{itemize}
	\item In the process mining stage, instead of \textit{Inductive Miner}, different techniques can be used which can mine complex process models with higher appropriateness levels while keeping the current high fitness values.
	\item In the performance indicator analysis stage, new indicators can be defined based on the business environment, event log attributes and user needs. For instance, personnel and resource allocation indicators can be included as well as cost dimension.
	\item For mismatch pattern analysis, new and business oriented mismatch patterns can be included in the analysis. In addition analyzers can fail when there are loops in the process models in current implementations, therefore more robust implementations for process models with loops can be developed in the future.
	\item For the generated recommendations, their quality for business environment is not assessed within the scope of this thesis. However, when any feedback from a domain expert or BPM people is provided, the learning approach can be converted to semi-supervised learning from unsupervised learning.
\end{itemize}

>>>>

In this study, a new approach is proposed and tested for generating recommendations using cross-organizational process mining for process performance improvement. Cross-organizational process mining is applied with the idea of unsupervised learning where predictor variables related to performances of organizations are used in an environment where processes are executed on several organizations. Results show that it is possible to use cross-organizational process mining and mismatch patterns for performance improvement recommendations. In this study, a four-stage solution is presented and their performances are explained. In addition, proposed methodology is developed as extensible and configurable set of plugins in ProM framework \cite{verbeek2010prom} and published as open-source. This makes the methodology open to include new process mining methods, mismatch patterns and clustering approaches as well as testing with different datasets.

For the approach proposed in this study, the following issues can be listed as pointers to future work:
\begin{itemize}
	\item In the process mining stage, instead of \textit{Inductive Miner}, different techniques can be used which can mine complex process models with higher appropriateness levels while keeping the current high fitness values.
	\item In the performance indicator analysis stage, new indicators can be defined based on the business environment, event log attributes and user needs. For instance, personnel and resource allocation indicators can be included as well as cost dimension.
	\item For mismatch pattern analysis, new and business oriented mismatch patterns can be included in the analysis. In addition analyzers can fail when there are loops in the process models in current implementations, therefore more robust implementations for process models with loops can be developed in the future.
	\item For the generated recommendations, quality for business environment is not assessed within the scope of this study. However, when any feedback from a domain expert or BPM people is provided, the learning approach can be converted to semi-supervised learning from unsupervised learning.
\end{itemize}
